### 2017-10-20 Writing/draft.tex
Removed capital letters for character difference

\newpage
\begin{abstract}
- When using discrete morphological characters for building phylogenies, the characters are assumed to be independent.

- However, this assumption is nearly always wrong, simply because of the nature of morphological integration in organisms.

- The dependence between characters can originate from biological features, shared phylogenetic history or errors from the researcher when coding characters.

- Although the two first sources can be investigated by biologists post phylogenetic inference and the third one can be avoided pre inference with good practices, software do not distinguish between any of them.

- In this study, we propose a new metric of raw character difference as a proxy for character (in)dependence.

- Using thorough simulations and this metric, we test the effect of increasing or decreasing Character Difference on topology with a variable number of taxa and characters under both Bayesian and Maximum Parsimony inference.

- Overall we found a positive effect of reducing character correlations on recovering the correct topology.

- However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation was not more effective than randomly drawing characters.

- Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.

- These results suggest that ignoring the problem of character correlation or independence can easily impact topology in phylogenetic analysis.

- However, encouragingly, they also suggest that, unless correlation is effectively maximised or minimised, probabilistic methods can easily accommodate for random correlation between characters.

@@

\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
The three types sources are however dissociated by later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.

@@

We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
\begin{enumerate}
    \item \textbf{Simulating matrices}: we simulated a discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices. This step resulted in 9 matrices.
    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising Character Difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy). This step resulted in 36 matrices.
    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees). This step resulted in 72 topologies.
    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
\end{enumerate}
Each step was replicated 40 times and are described below in more details, as well as our proposed definition for measuring the difference between characters.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a Character Difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration Supplementary Material 1).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
We standardised each character by arbitrarily modifying their character state tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).

When the Character Difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
When the Character Difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
It is important to stress that a Character Difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different phylogenetic interpretation.

@@

\subsection{Modifying the matrices}
We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
We then modified the matrices to either maximise or minimise the pairwise Character Differences for each matrices using three different algorithms.
For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e. with an average Character Difference $<0.25$) and replaced them randomly by any of the remaining characters.
This operation increased the overall pairwise Character Difference in the matrix thus making the characters more dissimilar.
Conversely, for minimising the pairwise Character Differences, we selected the most dissimilar characters (i.e. with an average Character Difference $<0.75$) and randomly replaced them with the remaining ones.
Finally, because this operation effectively changes the weight of characters (i.e. giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e. randomly weighting characters).
Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
The algorithms for the three modifications are available on GitHub@@@. %(Modifying_matrices.html) %TG: "@@@" is my "to replace" sign.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise Character Differences for 100 characters. Blue colours correspond to low Character Differences and orange colours correspond to high Character Differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised Character Difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
\label{Fig:modif_matrix}
\end{figure}

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
\caption{\small{Effect of Character Difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

\subsection{Effect of Character Differences on topology}

The overall amount of Character Difference in a matrix has an effect of the ability to recover the correct topology with maximising the Character Difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the Character Difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the Character Difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the Character Difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})

@@

This effect of the Character Difference is not dependent on the number of characters for the clade conservation (i.e. $NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the Character Difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.

@@

Similarly than the effect of number of characters on Character Difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

\subsection{Effect of Character Differences on the inference method}

@@

\subsection{Effect of Character Differences on topology}
As expected, there is an significant effect of the characters difference in the ability to recover the correct topology.
The Character Difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high Character Difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every characters agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
Intuitively, this would lead to the ``minimised'' Character Difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).

Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the Character Differences.
Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e. as an equivalent of randomising the Character Differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).

@@

Although we measured a significant difference between ``small'' a larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
For the number of characters, this suggests that the overall Character Difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
Similarly, using the a Normalised Tree Similarity metric ($NTS$) allows to account for the fact that topological difference is affected by the sheer number of taxa considered (i.e. we here corrected for the expected difference when comparing two random trees with the same number of taxa).

\subsection{Effect of Character Differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e. those with in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low Character Differences.
In fact the differences observed here could be due to the inherent mechanisms of each method.
When many characters are homoplasic (high character difference) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
These two points can explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.

@@

We found that Character Differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low Character Differences), the topology was always the furthest away from the ``normal'' topology.
Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
These results seems independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.

>>>>

\newpage
\begin{abstract}
Morphological characters for building phylogenies are assumed to be independent.
However, this assumption is nearly always wrong simply because of the nature of morphological integration of organisms.
Correlation between characters can originate from biological features, shared phylogenetic history or errors when coding or defining morphological characters.
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.

In this study, we propose a new metric of raw character difference as a proxy for character correlation.
Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
Overall, we found an expected positive effect of reducing character correlations on recovering the correct topology.
However, this effect is less important for matrices with a small number of taxa (~25) were reducing character correlation is not more effective than randomly drawing characters.
Furthermore, in bigger matrices (~350 taxa), there is a strong effect of the inference method with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.

These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.

@@

\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!
The three types sources are however dissociated by the later (coding dependence) being detected prior to phylogenetic inference and the two other ones (biological and evolutionary) being only detected and/or tested \textit{a posteriori} based on a phylogenetic hypothesis.

@@

We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
\begin{enumerate}
    \item \textbf{Simulating matrices}: we simulated a discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices. This step resulted in 9 matrices.
    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by modified the characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character differences to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy). This step resulted in 36 matrices.
    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees). This step resulted in 72 topologies.
    \item \textbf{Comparing topologies}: finally, we compared the ``normal'' to the ``maximised'', ``minimised'' and ``randomised'' trees to measure the effect of character correlation on topology.
\end{enumerate}
Each step was replicated 40 times and are described below in more details, along with our proposed definition for measuring the difference between characters.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using Maximum Parsimony and Bayesian inference (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the Supplementary Material 1).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
We standardised each character by arbitrarily modifying their character state tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).

When the character difference null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated).
When the character difference maximal ($1$) it means it conveys the most different signal (i.e. characters are uncorrelated).
It is important to stress that a character difference of $0$ (i.e. the same phylogenetic signal) does not mean the opposite of $1$ (i.e. \textit{not} the opposite phylogenetic signal) .
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.
Because the character is continuous and bounded between $(0,1)$, it can be interpreted as the probability of two characters leading to a different phylogenetic interpretation.

@@

\subsection{Modifying the matrices}
We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
We then modified the matrices to either maximise or minimise the pairwise character Differences for each matrices using three different algorithms.
For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e. with an average character difference $<0.25$) and replaced them randomly by any of the remaining characters.
This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e. with an average character difference $<0.75$) and randomly replaced them with the remaining ones.
Finally, because this operation effectively changes the weight of characters (i.e. giving the characters $<0.25$ or $>0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e. randomly weighting characters).
Each of the three matrices are effectively a bootstrap pseudo-replication of the ``normal'' matrix with the ``randomised'' one being a random one and the ``maximised'' and ``minimised'' being conditional bootstraps.
This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
The algorithms for the three modifications are available on GitHub@@@. %(Modifying_matrices.html) %TG: "@@@" is my "to replace" sign.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
\label{Fig:modif_matrix}
\end{figure}

@@

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and Supplementary materials 3).

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
\caption{\small{Effect of character difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different character difference scenarios and tree inference method with the maximised character difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised character difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised character difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

\subsection{Effect of character differences on topology}

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology with maximising the character difference leading to the less loss in phylogenetic information (median $NTS_{RF}$ = 0.95 and median $NTS_{Tr}$ = 0.96) followed by simply randomising the characters (median $NTS_{RF}$ = 0.79 and median $NTS_{Tr}$ = 0.72) and minimising the character difference (median $NTS_{RF}$ = 0.6 and median $NTS_{Tr}$ = 0.28 - see supplementary material 3 for the full summary statistics).
There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya coefficient of 0.89 for the $NTS_{RF}$ and 0.92 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.59 for the $NTS_{RF}$ and 0.63 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})

@@

This effect of the character difference is not dependent on the number of characters for the clade conservation (i.e. $NTS_{RF}$) for a ``similar'' number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.76, 0.78, 0.8 respectively - see supplementary materials 3) with a significant effect only between 100 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxa placement for a low number of characters (median $NTS_{RF}$ for 100, 350 and 1000 characters equals 0.54, 0.70, 0.76 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
These difference have to be contrasted however by a really high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.

@@

Similarly than the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences in both $NTS_{RF}$ and $NTS_{Tr}$ (medians for 25, 75 and 150 taxa equals 0.8, 0.76, 0.75 $NTS_{RF}$ and 0.75, 0.59 and 0.59 $NTS_{Tr}$respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

\subsection{Effect of character differences on the inference method}

@@

\subsection{Effect of character differences on topology}
As expected, there is an significant effect of the characters difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every characters agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasies.
Intuitively, this would lead to the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).

Following these results, we encourage researchers to code characters that appear to be the most un-correlated \textit{a priori}.
However, it is noteworthy to point that in rather small matrices, there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
Since many matrices are in this size range \citep{guillerme2016assessment} a simple bootstrap analysis (i.e. as an equivalent of randomising the character differences) can be sufficient to obtain good results (\textit{cf.} actively collecting different characters).

@@

Although we measured a significant difference between ``small'' a larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation and is independent of the number of characters analysed.
Similarly, using the a Normalised Tree Similarity metric ($NTS$) allows to account for the fact that topological difference is affected by the sheer number of taxa considered (i.e. we here corrected for the expected difference when comparing two random trees with the same number of taxa).

\subsection{Effect of character differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the Maximum Parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of both methods in recovering the ``right'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will more readily recover the ``right'' topology when analysing discrete morphological alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have been made in favour of parsimony: \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e. those with in a $CI>0.26$ in a quick parsimony search;]{OReilly20160081} could also favour Parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

This being said, it is interesting to analysis this difference in tree inference method in the light of the effect of high/low character differences.
In fact the differences observed here could be due to the inherent mechanisms of each method.
When many characters are homoplasic (high character difference) parsimony will cope with this information by generating longer trees (which will be more likely to be discarded in the tree search).
Conversely, Bayesian analysis can (1) cope with homoplasy more easily by increasing branch length to accommodate more changes and (2) will also incorporate sub-optimal trees in the final results.
These two points can explain the fact that the effect of character correlation seems less important in Bayesian inference, especially in terms of clade conservation.

@@

We found that character differences as a proxy for character correlation have a strong effect on the recovering the ``normal'' topology: when character correlation was high (low character differences), the topology was always the furthest away from the ``normal'' topology.
Conversely, when correlation between characters was low, the topology was always the closest to the ``normal'' topology.
These results seems independent on the size of the matrix (number of taxa and/or characters) but can be influenced by the phylogenetic inference method used with Bayesian inference faring better in terms of clade conservation, especially in larger matrices.
